---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aihuman

<!-- badges: start -->
  [![R-CMD-check](https://github.com/sooahnshin/aihuman/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sooahnshin/aihuman/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Overview

`aihuman` provides statistical methods for analyzing experimental evaluation of the causal impacts of algorithmic recommendations on human decisions developed by [Imai, Jiang, Greiner, Halen, and Shin (2023)](https://doi.org/10.1093/jrsssa/qnad010).

The data used for this paper, and made available here, are interim, based on only half of the observations in the study and (for those observations) only half of the study follow-up period.  We use them only to illustrate methods, not to draw substantive conclusions.

## Installation

You can install the development version of `aihuman` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sooahnshin/aihuman", dependencies = TRUE, build_vignettes = TRUE)
```

If you have trouble with compilation on macOS, you may want to check this [link](https://mac.thecoatlessprofessor.com/macrtools/).

## Usage 1: Evaluation Based on Principal Stratification

The package provides main functions for the methods proposed by [Imai, Jiang, Greiner, Halen, and Shin (2023)](https://doi.org/10.1093/jrsssa/qnad010) based on principal stratification.

| Category | Function | Type | Main Input | Output | Paper | Notes |
|:---------|:---------|:-----|:------|:-------|:------|:------|
|descriptive|`PlotStackedBar()`|vis.|e.g., `data(psa_synth)`           |ggplot    |Fig 1 |dist. of $D_i$|
|descriptive|`CalDIMsubgroup()`   |est.|e.g., `data(synth)`               |dataframe |Sec 2.4|diff-in-means|
|descriptive|`PlotDIMdecisions()` |vis.|output of `CalDIMsubgroup()`      |ggplot    |Fig 2 (left)||
|descriptive|`PlotDIMoutcomes()`  |vis.|output of `CalDIMsubgroup()` for each outcome|ggplot    |Fig 2 (right)||
|main       |`AiEvalmcmc()`          |est.(c/h)|e.g., `data(synth)`          |mcmc      |Sec S5           ||
|main       |`CalAPCE()`          |est.(c/h)|output of `PSAmcmc()`        |list      |Sec 3.4||
|main       |`APCEsummary()`      |est.|output of `CalAPCE()`             |dataframe |Sec 3.4|APCE|
|main       |`PlotAPCE()`         |vis.|output of `APCEsummary()`         |ggplot    |Fig 4||
|strata     |`CalPS()`            |est.|`$P.R.mcmc` of `CalAPCE()`   |dataframe |Eq 6 |$e_r$|
|strata     |`PlotPS()`           |vis.|output of `CalPS()`               |ggplot    |Fig 3||
|fairness   |`CalFairness()`      |est.|output of `CalAPCE()`             |dataframe |Sec 3.6|$\Delta_r(z)$|
|fairness   |`PlotFairness()`     |est.|output of `CalFairness()`         |ggplot    |Fig 5||
|optimal    |`CalOptimalDecision()`|est.|output of `AiEvalmcmc()`            |dataframe |Sec 3.7|$\delta^\ast(\mathbf{x})$|
|optimal    |`PlotOptimalDecision()`|vis.|output of `CalOptimalDecision()`|ggplot    |Fig 6||
|comparison |`PlotUtilityDiff()`  |vis.|output of `CalOptimalDecision()`  |ggplot    |Fig 7|$g_d(\mathbf{x})$|
|comparison |`PlotUtilityDiffCI()`|vis.|output of `CalOptimalDecision()`  |ggplot    |Fig S17||
|crt        |`SpilloverCRT()`|est.|court event hearing date, $D_i$, $Z_i$  |daraframe    |Sec S3.1||
|crt        |`PlotSpilloverCRT()`|vis.|output of `SpilloverCRT()`  |ggplot    |Fig S8||
|crt power  |`SpilloverCRTpower()`|est.|court event hearing date, $D_i$, $Z_i$  |dataframe    |Sec S3.2||
|crt power  |`PlotSpilloverCRTpower()`|vis.|output of `SpilloverCRTpower()`  |ggplot    |Fig S9||
|frequentist |`CalAPCEipw()`|est.| e.g., `data(synth)` |dataframe |Sec S7||
|frequentist |`BootstrapAPCEipw()`|est.| e.g., `data(synth)` |dataframe |||
|frequentist |`APCEsummaryipw()`|est.| outputs of `CalAPCEipw()` and `BootstrapAPCEipw()` |dataframe |||
|frequentist (RE) |`CalAPCEipwRE()`|est.| e.g., `data(synth)` |dataframe |||
|frequentist (RE) |`BootstrapAPCEipwRE()`|est.| e.g., `data(synth)` |dataframe |||

vis. = visualization; est. = estimation; c/h = computation-heavy.
You may use `CalAPCEparallel()` instead of `CalAPCE()` throughout the analysis.

For more details, see the `aihuman` package vignette, `vignette("aihuman", package = "aihuman")`.

### Example

```{r example}
library(aihuman)
## Using synthetic data with small run
data(synth)
sample_mcmc <- AiEvalmcmc(data = synth, n.mcmc = 10)
subgroup_synth <- list(1:nrow(synth),
                       which(synth$Sex == 0),
                       which(synth$Sex == 1),
                       which(synth$Sex == 1 & synth$White == 0),
                       which(synth$Sex == 1 & synth$White == 1))
sample_apce <- CalAPCE(data = synth, 
                       mcmc.re = sample_mcmc, 
                       subgroup = subgroup_synth)
# You can also use the parallelized version: check CalAPCEparallel()
sample_apce_summary <- APCEsummary(sample_apce[["APCE.mcmc"]])
PlotAPCE(sample_apce_summary, 
         y.max = 0.25, 
         decision.labels = c("signature", "small cash", "medium cash", "large cash"), 
         shape.values = c(16, 17, 15, 18), 
         col.values = c("blue", "black", "red", "brown", "purple"), 
         label = FALSE)
```

## Usage 2: Evaluation With Minimal Assumptions

The package provides main functions for the methods proposed by [Ben-Michael, Greiner, Huang, Imai, Jiang, and Shin (2024)](https://arxiv.org/pdf/2403.12108) based on a minimal set of assumptions.

| Category | Function | Type | Output | Figure \# |
|:---------|:---------|:-----|:------|:-------|
|Human+AI v. Human|`compute_stats_aipw()`|est.|dataframe    | |
|Human+AI v. Human|`plot_diff_human_aipw()`|vis.|ggplot    |Fig 1 |
| AI v. Human|`compute_bounds_aipw()`|est.|dataframe    | |
| AI v. Human|`plot_diff_ai_aipw()`|vis.|ggplot    |Fig 2, 5, S4 |
| Preference|`plot_preference()`|vis.|ggplot    |Fig 3, S5 |
| Agreement|`table_agreement()`|est|ggplot    | |
| Agreement|`plot_agreement()`|vis.|ggplot    |Fig S1 |
| Overrides|`plot_diff_subgroup()`|vis.|ggplot    |Fig S2, S3 |
| Policy learning|See the vignette|vis.|ggplot    |Fig 4 |

For more details, see the `ablity` vignette, `vignette("ability", package = "aihuman")`.

### Example

```{r example2}
library(ggplot2)
## set default ggplot theme
theme_set(theme_bw(base_size = 15) + theme(plot.title = element_text(hjust = 0.5)))

## Human+AI v. Human
plot_diff_human_aipw(
  Y = NCAdata$Y,
  D = ifelse(NCAdata$D == 0, 0, 1),
  Z = NCAdata$Z,
  nuis_funcs = nuis_func,
  true.pscore = rep(0.5, nrow(NCAdata)),
  l01 = 1,
  subgroup1 = ifelse(NCAdata$White == 1, "White", "Non-white"),
  subgroup2 = ifelse(NCAdata$Sex == 1, "Male", "Female"),
  label.subgroup1 = "Race",
  label.subgroup2 = "Gender",
  x.order = c("Overall", "Non-white", "White", "Female", "Male"),
  p.title = NULL, p.lb = -0.3, p.ub = 0.3
)

## AI v. Human
plot_diff_ai_aipw(
  Y = NCAdata$Y,
  D = ifelse(NCAdata$D == 0, 0, 1),
  Z = NCAdata$Z,
  A = PSAdata$DMF,
  z_compare = 0,
  nuis_funcs = nuis_func,
  nuis_funcs_ai = nuis_func_ai,
  true.pscore = rep(0.5, nrow(NCAdata)),
  l01 = 1,
  subgroup1 = ifelse(NCAdata$White == 1, "White", "Non-white"),
  subgroup2 = ifelse(NCAdata$Sex == 1, "Male", "Female"),
  label.subgroup1 = "Race",
  label.subgroup2 = "Gender",
  x.order = c("Overall", "Non-white", "White", "Female", "Male"),
  zero.line = TRUE, arrows = TRUE, y.min = -Inf,
  p.title = NULL, p.lb = -0.3, p.ub = 0.3
)
```

